### 2018-04-20 RefereeReportResponse-17Apr2018.txt
Start addressing referee comments

----------------------------------------------------------------------
Report of Referee A -- LQ16046/Thyagarajan
----------------------------------------------------------------------

In this paper, the authors propose using the bispectrum phase (i.e.,
closure phase in radio interferometry) as a way to detect the 21cm
signal from cosmic reionization. This is attractive because
antenna-based gain factors naturally cancel out of the observable
quantities, so a reported detection of the cosmological signal would
be more robust to systematics. Given the difficulties in calibrating
low-frequency interferometers to high precision, I think this work is
extremely important. I would be happy to recommend acceptance after my
comments below are addressed. This paper is suitable for publication
in PRL rather than (say) PRD because for years the 21cm community has
searched for an alternative to the power spectrum. The results are
either not scientifically interesting or are not observationally
realistic. This paper is a rare example that does not suffer from
these shortcomings.

Specific comments:

1) Introduction: the comment about the photon-to-baryon ratio is a
rather strange comment, because the statement is only true if one
ignores the CMB in calculating the ratio.

[Response] Modified to "photon-to-baryon ratio rose significantly"

2) Penultimate paragraph of introduction: I think it's debatable
whether current experiments have already "obtained sufficient
observational data"

[Response] Modified "sufficient" to "significant"

3) The summation/product notation should be revised to conform to
convention. Generally, if one writes something like \sum_{ij}, I
interpret this to mean the double sum \sum_{i=1}^N \sum_{j=1}^N. With
N = 3 antennas in this paper, I would thus interpret the sums to give
rise to 9 terms, including the i = j terms. Instead, I think the
authors mean a single sum over 1-2, 2-3, 3-1.

[Response] The referee is correct. We mean a single sum over 1-2, 2-3,
and 3-1.

4) I am not sure why the equation for B_{\Del}^m does not contain more
terms. Each V_ij has the form (signal + noise), so when I multiply
three such quantities together I should get terms that look like
signal^2 noise and signal noise^2. The signal^2 noise term has zero
mean if the noise on a visibility has zero mean, while the signal
noise^2 term will have zero mean if the different baselines have
independent noise. So I could believe that they're zero mean, but then
so is the pure noise term in the equation, and so I don't understand
why only that pure noise term was included.

[Response] Due to constraints on space, we include all cross-terms and
noise-only terms in B_\nabla^N. Thus it comprises of all terms including
cross-terms that contain visibility noise. And in our analysis, the
triads are made of antennas i \ne j \ne k. Thus B_\nabla^N is zero-mean.

5) On a different, but related note, I think that the noise term
B_{\Del}^N only has mean zero if one assumes that the noise
contributions from different baselines are uncorrelated. It's not
sufficient to just say that it has zero mean because it contains V^N.

[Response] The referee is correct. We assume the noise across baselines
is uncorrelated.

6) Ultimately, I felt that the authors should probably have discussed
a little more whether in practice one would expect a) negligible gain
factors that cannot be parameterized on a per-antenna basis, and b)
noise that is necessarily independent between baselines.

[Response] In practice, both these items can only be evaluated from
data. Since these non-ideal effects are instrument specific, and
difficult to estimate without data, we have left it for the data
analysis paper. We will present the results of such data analysis from
HERA in a forthcoming paper. For purposes of concept demonstration, in
this paper we have assumed non-antenna based gains are negligible, and
noise is uncorrelated across baselines.

7) I think the authors should clarify that \delta \phi_{ij}^N is *not*
the phase of B_{\Del}^N, lest the unsuspecting reader think otherwise.
My understanding is that taking the natural logarithm of a complex
number to get at the phase, we have something like ln(s+n) = ln s + ln
(1 + n / s) ~ ln s + n/s. So the phase \delta \phi_{ij}^N is
noise-to-signal ratio itself, rather than the phase of the noise. Is
that correct?

[Response] That is correct. In fact, we have stated the same fact in
the first sentence of last paragraph of section II. The referee's
point about the unsuspecting reader is well taken. The noise in the
visibility measurement translates to phase fluctuations (relative to
true sky phase) whose magnitude (or RMS) is a measure of 1/(S/N).
The phase fluctuations caused by noise are NOT the phase of the noise.
It also depends on the sky signal strength. It is only a notation we
have employed, but it has to be interpreted as not the phase of the
noise itself but fluctuations to phase caused by presence of noise
and these fluctuations \delta\phi_{ij}^N are perturbations over the
true phase and have an RMS 1/(S/N). We have modified text below
Equation 1 to clarify this point.

8) In the final paragraph of Section II, it seems like the authors are
ignoring variance from the foregrounds. Some foregrounds (e.g.
extremely bright point sources) would be inappropriate to describe in
terms of a variance, but for others (e.g., dim, unresolved point
sources), a statistical distribution like a Poisson distribution would
be an appropriate description. That would certainly contribute to the
variance.

9) Note that the previous point also affects comment 7) above. If
\delta \phi_{ij}^N goes like n/s, then yes, if n is Gaussian and s is
some definite signal with no variance, then n/s will be Gaussian. But
if s has its own distribution, then n/s may not be Gaussian. I'm not
asking for a full treatment of all probability distributions, but
(assuming that I'm right about these points), some discussions of the
caveats would be important.

10) In the 3rd paragraph of Section IV, shouldn't it be |V_F| rather
than V_F? At least that's what I would've thought given the definition
of rho.

[Response] Correct. Fixed now.

11) When the authors define the cross-power spectrum, it would be
worth mentioning which sorts of axes they imagine computing these
cross spectra.

[Response] This is discussed in Section IV B. Cross-power spectrum
would be used in cases where redundancy cannot be assumed. In section
IV B, incoherent averaging is done over corss-power spectrum. For
example, the sky is redundant in LST, that is it repeats every day at
a given LST. In this case, the closure phases themselves can be averaged
coherently before computing the power spectrum. Similarly, if the
baselines are truly redundant, those closure phases can be coherently
averaged. But if such a redundancy in baselines or neighboring LST scans
cannot be assumed, one would have to perform incoherent averaging on
cross-power spectra along axes which lack redundancy. The assumption
behind incoherent averaging of cross-power spectra is that the
cosmological signal is statistically isotropic in space whereas the
foregrounds and noise measurements can be averaged down (though not
as fast as coherent averaging).

12) I am curious: does Equation 3 vanish (in expectation) if I try to
form a cross-power spectrum with two triads that share one edge?

[Response] If the two triads are not redundant but share an edge, it
depends on whether the measurement is sky noise dominated or receiver
noise dominated. The former is generally the case at low frequencies.
In such a scenario, there will be a non-zero correlation between the
measurements on the two triads and thus the expectation of equation 3
will not necessarily vanish. Even when the measurements are receiver
noise dominated, and sufficient coherent averaging has been performed
to an extent that receiver noise contributions have been rendered
negligible, the correlations from residual sky noise may then be
visible between the two triads.

13) In Section IV C, "HERA-47" is not defined. I assume this is HERA
with 47 dishes, but how are they arranged in this intermediate stage
of HERA's buildout?

[Response] The HERA-47 layout has been illustrated in the companion
paper [Carilli et al. 2018]. We are unable to describe further due to
lack of space but have modified appropriately with a citation to the
companion paper.

14) Figure 3 needs to be made considerably easier to read. For one, it
is the main result of the paper and deserves to be a much bigger plot.
(This is not because I'm old and have bad eyesight...let's just say
that I am within 10 years of the age of the lead author of this
paper). Some of the colors are quite similar, and I would also like to
be able to know what the different curves are plotting without having
to read the caption.

15) The signal-to-noise ratio (inferred from Figure 3) seems to be
quite high even at high k. Does this trend continue beyond the values
plotted? If so, this is a very interesting property of the bispectrum
phase, because the regular power spectrum loses signal to noise
quickly with k. I wonder if this is a promising way to get at some
high k information? I guess it's hard to tell because it's not clear
what the intuition is like in terms of the scattering of power over
different k's.

16) Do the conclusions of the last paragraph of Section II require
assuming that the HI fluctuations are Gaussian? (Which of course they
aren't, although I'm fine with that assumption being made for
simplicity for this paper).

----------------------------------------------------------------------
Report of Referee B -- LQ16046/Thyagarajan
----------------------------------------------------------------------

I feel this paper should be published in PRL. It proposes an elegant
solution to a challenge that has plagued interferometers attempting to
exploit the faint radio signal from neutral hydrogen gas at high
redshifts as a cosmological tool for understanding the epoch of
reionization.

Indeed, my main suggestion is to expand the scope of the paper by
mentioning in the abstract and introduction that the technique has
broader application than just to measurements of the EoR. I believe
this broader focus would increase the impact of the paper
significantly. Cosmological measurements of HI at lower redshifts (<
2) are also being pursued vigorously by a variety of radio
interferometers (CHIME, Tianlai, HIRAX) that face the same challenges
as EoR instruments such as HERA. The simulations described in the
paper using HERA as an example are equally relevant for these other
interferometers, which also must precisely calibrate instrumental
gains in the presence of strong astrophysical foregrounds. There is
considerable interest in the cosmological community to develop HI
intensity mapping as an efficient tool for mapping large-scale
structure. The hoped-for payoff is improved precision on cosmological
parameters such as the neutrino mass and the equation of state of the
dark energy. Outside of the introduction, I believe that little or
nothing would need to be changed in the rest of the paper.

I do have one minor suggestion:

Just after equation 1, a phrase should be reversed to read “small
perturbations due to EoR HI signal and noise, respectively”.

[Response] Fixed.