### 2018-04-27 ResponseToEditor.txt
Update on Overleaf.

To the editorial staff:

We appreciate the referees' reports and their positive views on 
our work. We have responded to their comments, attached in a 
separate file.

We are attaching two versions of the Latex file. One is the 
updated ms.tex (ms.pdf). The other is ms_changes.tex (and 
ms_changes.pdf) and it contain all the changes highlighted in 
red.

Regarding the length of the manuscript, we believe there has 
been a mistake. In the email from PRL on 7 March 2018, the word 
count summary was stated to be:

-----------------------------------------------------------------
  Figure   Aspect Ratio   Wide?   Word Equivalent
     1         1.54         No        117
     2         1.06         No        161
     3         3.73        Yes        200

WORD COUNT SUMMARY
    Note: Text word count excludes title, abstract, byline, PACS,
          receipt date, acknowledgments, and references.

              Text word count   4003
        Equations word equiv.    192
          Figures word equiv.    478
           Tables word equiv.      0
                              ------
                       TOTAL    4673 (Maximum length is 3750)
-----------------------------------------------------------------

Subsequently, we made substantial changes to reduce the length by
at least 100 words including substantial changes to figure sizes. 
We confirmed this through "texcount" on the command line and 
carefully kept track of the word count at the time of initial 
submission (5 March 2018) as well as the trimmed one (12 March 
2018).

In the recent email from PRL on 17 April 2018, the word count 
was found by PRL to be: 
-----------------------------------------------------------------
  Figure   Aspect Ratio   Wide?   Word Equivalent
     1         1.54         No        117
     2         1.06         No        161
     3         3.73        Yes        200

WORD COUNT SUMMARY
    Note: Text word count excludes title, abstract, byline, PACS,
          receipt date, acknowledgments, and references.

              Text word count   4003
        Equations word equiv.    192
          Figures word equiv.    478
           Tables word equiv.      0
                              ------
                       TOTAL    4516 (Maximum length is 3750)
-----------------------------------------------------------------
You will notice that the word count for the text, equations, and 
figures are exactly identical between 17 April and 7 March (which
is impossible), although the total has changed from 4673 to 4516. 
Not only has the itemized count not been updated, but the new 
total is also inaccurate.

We understand the output of texcount may not be identical to the 
software PRL uses. Therefore, we request PRL to give us an 
accurate count which we can use as a reference to further trim 
the manuscript, if necessary.

In the meanwhile, we request that the editorial office pass on
the revised manuscript to the referees for a speedy acceptance.

Best regards,
The authors.